---
title: "Data Summary and Analysis"
author: "Robert Liddell"
date: "2022-10-27"
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
output: html_document
---



```{r Setup-Libraries Analysis Loading, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Code\\dataAnalysisCode.R")
```


```{r Formatting Functions, include=FALSE}
#Standardized table format
tableFormat<-function(tib,capt=""){
  tib %>% 
    kbl(caption=capt) %>% 
    kable_paper(full_width = T,"striped")
}

##Specify Default position_dodge characteristics
pDodge<-position_dodge(width=0.3)

##Figure Theme
theme_set(theme_minimal()+theme(plot.caption = element_text(hjust = 0)))
impCols <- scale_color_manual(values=c("black","grey70"))
impFills <- scale_fill_manual(values=c("black","grey70"))


##Plots the 95% confidence Interval with dat which specify term, upperCI and lowerCI
#plotting function for 95% confidence interval
term95ConfIntervalPlot<-function(dat){
  p<-dat %>%
    mutate(colour=if_else(CIupper*CIlower>0,'red','black')) %>% 
    ggplot() +
    geom_errorbar(aes(x=term,ymax=CIupper,ymin=CIlower,colour=colour), width=0.25) +
    geom_hline(yintercept=0,lty=3)+
    scale_colour_identity()+
    coord_flip()
  return(p)
}


```

# Goal
Summarize the data collected and perform statisitical analysis on results.

## Data Summary

The study was designed to place 240 implants (120/implant group) into 120 rats. with each implant being tested twice due to the lateral and medial arches this would result in 480 different test files if all samples were tested. These 480 different tests would be split between the 2 testing methods (Shear/Tension), 2 implant types (BAE/BAE+DCD), 2 arches (Lateral/Medial), 2 animal sides (Left/Right) and 6 time points (5/9/14/28/84/168 days). To simplify analysis this study will use the mean of the two arches, dropping samples which do not have both arches.

```{r Data Summary-Counts}
groupCounts %>% 
  tableFormat("Table 1. N for each implant type and testing method")
```
The graph below shows the distribution of data from the different implant groups, testing methods, and time points. It is evident from this figure that the range of data gets larger with longer time points as the median anchorage increases. It is also evident that BAE+DCD implants have greater anchorage than the BAE implants, particularly at early time points.Lastly one can also see the heteroskedastic nature of the data with early time points having smaller deviation than later time points.

```{r Data Summary-Box Plots}

pDodge<-position_dodge(width=0.3)

simpDat%>%
  group_by(testMethod,time,implantType) %>%
  summarise(Quarts=list(quantile(force)),
            .groups="drop") %>% 
  unnest_wider(Quarts) %>% 
  ggplot(aes(x=factor(time),colour=implantType))+
    geom_hline(aes(yintercept=0))+
    geom_linerange(aes(ymin=`0%`,ymax=`100%`),linewidth=1,position = pDodge)+
    geom_linerange(aes(ymin= `25%`,ymax=`75%`),linewidth=3,position = pDodge)+
    geom_point(aes(y=`50%`),size=6,shape='_',position = pDodge)+
    facet_grid(rows=vars(testMethod))+
    labs(x="Time (d)",y="Torque (Ncm)",title="Distribution of Data",caption=str_wrap("Figure 1. The transitions indicate 0%, 25%, 50%, 75%, and 100% quartiles",width=100))+
    impCols

```
## Growth Curves

The growth rates show that there is an initial peak between 5 and 9 days, after which anchorage slows down significantly. While the trends between the two testing methods are similar the BAE+DCD implant has a higher peak growth in anchorage compared to the BAE implant

```{r Growth Rates-% of Max}

simpDat%>%
  group_by(testMethod,time,implantType) %>%
  summarise(force=median(force),.groups = "drop") %>%
  add_row(expand_grid(testMethod=factor(c("Tension","Shear")),time=0,implantType=factor(c("BAE","BAE+DCD")),force=0)) %>% 
  group_by(testMethod,implantType) %>% 
  arrange(time) %>% 
  mutate(nforce=(force/max(force))*100,
         change=(nforce-lag(nforce))/(time-lag(time)),
         time=(time+lag(time))/2, .keep="unused") %>% 
  drop_na() %>% 
  ggplot(aes(x=factor(time),y=change,colour=implantType,group=implantType))+
  geom_line(linewidth=1)+
    facet_grid(rows=vars(testMethod))+
    labs(x="Time (d)",y="Growth Rate (%/Day)",title="Growth Over Time",caption=str_wrap("Figure 2. The rate of change in force between each time point for each implant. The percent change per day indicates how much growth occured relative to the peak measured anchorage.",width=100))+
    impCols+
  geom_hline(yintercept = 0)
  
```

```{r plotting Confidence Interval of Graph}

temp <- bootStrap_results_base|>
  filter(!startsWith(term,'D'))|>
  mutate(implantType=if_else(str_detect(term,'n'),'BAE+DCD','BAE'),
         testMethod=if_else(str_detect(term,'s'),'Shear','Tension'),
         term=str_sub(term,end=1))|>
  pivot_wider(id_cols=c(implantType,testMethod),
              names_from = term,
              values_from = c(Estimate,CIupper,CIlower))

bootStrap_base_fitCI %>% 
  ggplot(aes(x=time, fill=implantType)) +
  facet_grid(cols=vars(testMethod))+
  geom_ribbon(aes(ymin=`2.5%`,ymax=`97.5%`),alpha=0.7)+
  geom_line(aes(y=fit))+
  # geom_point(data=nlsDat,aes(x=time,y=force,fill=implantType),shape=21,colour='black')+
  geom_hline(yintercept=0)+
  geom_vline(xintercept=0)+
  geom_hline(data=temp,aes(yintercept=exp(Estimate_C),colour=implantType),lty=2,linewidth=1)+
  geom_linerange(data=temp,aes(ymin=0,ymax=exp(Estimate_C),x=exp(Estimate_T),colour=implantType),lty=3,linewidth=1)+
  impFills+
  impCols+
  labs(x="Time (d)",y="Torque (Ncm)",title="Asymptotic curve fit to Implant Anchorage", caption=str_wrap("Figure 3. Asymptotic curves fit to reverse torque data.",widht=100))
  


```



```{r Parameter Comparison}
bootStrap_results|>
  filter(str_detect(term,'[nt]$'))|>
  mutate(Nano=endsWith(term,'n'),
         Tension=endsWith(term,'t'),
         term=str_sub(term,end=1))|>
  pivot_wider(id_cols = c(Nano,Tension),
              names_from = term,
              values_from = c(Estimate,CIlower,CIupper))|>
  mutate(Effect=if_else(Nano,'Nano','Tension'))|>
  select(-Nano,-Tension)|>
  ggplot()+
  geom_hline(yintercept=1)+
  geom_vline(xintercept=1)+
  geom_errorbar(aes(x=exp(Estimate_T),ymax=exp(CIupper_C),ymin=exp(CIlower_C)),width=0.03)+
  geom_errorbar(aes(y=exp(Estimate_C),xmax=exp(CIupper_T),xmin=exp(CIlower_T)),width=0.05)+
  geom_point(aes(x=exp(Estimate_T),y=exp(Estimate_C),colour=Effect),size=3)+
  impCols+
  labs(x="Tau (d)",y="C (Ncm)",title="Relative Effect of testing method and Implant Surface", caption=str_wrap("Figure 4. 95% CI for the relative multiplicative effect of Nanotopography and Tensile testing method. CI's which do no overlap 1 result in a significant increase or decrease to the base parameter. Interactions between nanotopography and tension testing were not different from 1.", width=100))



```



